---
title: "Summary statistics"
subtitle: "011_Summary_stats.qmd"
format: html
eval: TRUE
warning: false
knitr: 
  opts_chunk:
    message: false
---

# General overview

In this manuscript we will provide the code needed to reproduce summary statistics reported in the *Results* section of the main text of the study *"The hidden biodiversity knowledge split in biological collections"*

Here you will find how we obtained the values of

-   Total fish binomial

-   Total fish binomials in countries of the Global South

-   Total fish binomials in countries of the Global North

-   Total name bearers in museums of Global South since 2000

-   Total name bearers in museum of Global North since 2000

-   Name bearers in Global North museums from Global South since 2000

-   Name bearers in Global South museums from Global North since 2000

In these descriptive statistics, we classified countries in East Asia and the Pacific, Europe and Central Asia, and North America as the Global North, while Latin America and the Caribbean, the Middle East and North Africa, South Asia, and Sub-Saharan Africa were considered the Global South.

# Packages and data used

Packages

```{r}
library(dplyr)   # data manipulation
library(tidyr)   # data tidying
library(glue)    # string interpolation
library(here)    # constructing file paths
library(readr)   # reading CSV files

# create period of 50 years
floor_period = function(value){ return(value - value %% 50) }
```

Data and data preparation

```{r}
flow_period_region_country <- readr::read_csv(here::here("data", "raw", "flow_period_region_country.csv"))
```

We first classified, according to the regions defined above, the countries in which the specimen has been collected (`region_type`) and the country of the museum in which it is housed was classified as Global North and Global South

```{r}
flow_period_region_country2 <- 
  flow_period_region_country |> 
  mutate(north.south.type = ifelse(region_type == "Latin America & Caribbean" | 
                                     region_type == "Sub-Saharan Africa" |
                                     region_type == "Middle East & North Africa" |
                                     region_type == "South Asia", "Global.South", "Global.North")) |> 
  mutate(north.south.museum = ifelse(region_museum == "Latin America & Caribbean" | 
                                       region_museum == "Sub-Saharan Africa" |
                                       region_museum == "Middle East & North Africa" |
                                       region_museum == "South Asia", "Global.South", "Global.North"))

```

# Total fish binomial

```{r}
total_fish_binomial <- 
  flow_period_region_country2 |> 
  pull(n) |> 
  sum()
```

The total fish binomial is `r total_fish_binomial`


# Total fish binomials in museum countries of the Global South

```{r}
total_fish_binomial_museum_south <- 
  flow_period_region_country2 |> 
  filter(north.south.museum == "Global.South") |> 
  pull(n) |> 
  sum()

```
The total fish binomial in museums of Global South is
    `r total_fish_binomial_museum_south`

# Total fish binomials in museum countries of the Global North

```{r}
total_fish_binomial_museum_north <- 
  flow_period_region_country2 |> 
  filter(north.south.museum == "Global.North") |> 
  pull(n) |> 
  sum()

```
The total fish binomial in museums of Global South is
    `r total_fish_binomial_museum_north`

# Total name bearers in museums within Global North and South since 2000

```{r}

# Total in Global North 
sum_post2000_total_north <- 
  flow_period_region_country2 |> 
  filter(period >= 2000,
         north.south.museum == "Global.North") |> 
  pull(n) |> 
  sum()

# Total in Global South
sum_post2000_total_south <- 
  flow_period_region_country2 |> 
  filter(period >= 2000,
         north.south.museum == "Global.South") |> 
  pull(n) |> 
  sum()

```

# Name bearers in Global North museums from Global South since 2000

Calculating the number of name bearers collected in Global South and housed in Global North since 2000

```{r}
sum_post2000_south_north <- 
  flow_period_region_country2 |> 
  filter(period >= 2000,
         north.south.type == "Global.South" & north.south.museum == "Global.North") |> 
  pull(n) |> 
  sum()
```
`r sum_post2000_south_north` name bearers housed in Global North biological 
    collections are from Global South. This represents 37% of the name bearers 
    in Global North since 2000s

# Name bearers in Global South museums from Global North since 2000

Calculating the number of name bearers collected in Global North countries and housed in Global South museums since 2000

```{r}
sum_post2000_north_south <- 
  flow_period_region_country2 |> 
  filter(period >= 2000,
         north.south.type == "Global.North" & north.south.museum == "Global.South") |> 
  pull(n) |> 
  sum()
```
Only `r sum_post2000_north_south` of name bearers housed in Global South 
    biological collections are from Global North


-   Proportion of name bearers from Global South housed in Global North museums since 2000 is of `r ((sum_post2000_south_north)/(sum_post2000_total_north)) *100`

- Proportion of name bearers from Global North housed in Global South museums since 
    2000 is of `r ((sum_post2000_north_south)/(sum_post2000_total_south)) *100`
    
    
