---
title: "07 - DC and DR for each region per time (Figure 2)"
format: html
eval: FALSE
---

# Domestic contribution (DC) and Domestic retention (DR) for all countries

Reading libraries

```{r}
library(dplyr)    # Data manipulation
library(here)     # File path management
library(tidyr)    # Data tidying
library(ggplot2)  # Data visualization
library(scales)   # Scale transformations and labels
library(ggarrow)  # Adding arrows to ggplot
```


Here we combined two metrics, DC and DR by time and regions, to express the
    changes of NBTs distribution for each of these regions. 
    This is aimed to be the Figure 02 in the manuscript. The Figure 2 in the main
    text was further edited in inkscape to better acomodate the arrows and 
    include the quadrant information in the first plot.


```{r}
# Data from 01_C_data_preparation.qmd

flow_period_region_prop <- readr::read_csv(here::here("data", "processed", "flow_period_region_prop.csv"))
```

Add lag values for visualization

```{r}
flow_period_region_prop_lag <- flow_period_region_prop |>
  dplyr::group_by(region_type) |>
  dplyr::mutate(new_DC = dplyr::lag(prop_DC),
                new_DR = dplyr::lag(prop_DR)) |> 
  dplyr::ungroup()
```

## Figure 2 - Scatterplot

```{r}
flow_period_region_prop_lag |>
  dplyr::filter(dplyr::if_all(c(prop_DC, prop_DR, 
                                new_DC, new_DR), 
                              ~ . != 0 | is.na(.))) |> 
  ggplot(aes(x = prop_DC, y = prop_DR, fill = region_type))+
  geom_hline(yintercept = 0.5, linetype = "dashed")+
  geom_vline(xintercept = 0.5, linetype = "dashed")+
  geom_arrow_segment(
    aes(x = new_DC, xend = prop_DC,
        y = new_DR, yend = prop_DR),
    color = "grey",
    arrow_head = NULL,
    arrow_mid  = arrow_head_wings(offset = 30,
                                  inset = 60),
    resect_head = 2,
    resect_fins = 2
  )+
  geom_point(
    shape = 21,
    size = 2.5
  )+
  geom_point(
    aes(x = new_DC ,
        y = new_DR),
    alpha = 0.5,
    shape = 21,
    size = 2.5
  )+
  facet_wrap(.~period,axes = "all",
             axis.labels = "margins"
              )+
  scale_fill_manual(
    values = c(
      "Europe & Central Asia" = "#E64B35FF",
      "East Asia & Pacific" = "#4DBBD5FF",
      "North America" = "#3C5488FF",
      "South Asia" = "#B09C85FF",
      "Latin America & Caribbean" = "#00A087FF",
      "Sub-Saharan Africa" = "#F39B7FFF",
      "Middle East & North Africa" = "#8491B4FF"
    )
  )+
  scale_x_continuous(
    labels = scales::label_percent(),
    expand = expansion(mult = c(0.05, 0))
  )+
  scale_y_continuous(
    labels = scales::label_percent(),
    expand = expansion(mult = c(0.05, 0))
  )+
  theme_classic()+
  theme(
    strip.background = element_rect(fill = NA, color = NA),
    strip.text = element_text(face = "bold"),
    legend.position = "none",
    plot.background = element_blank(),
    panel.spacing = unit(5, "pt"),
    panel.spacing.x = unit(15, "pt"),
    plot.margin = margin(5,15,5,5,"pt"),
    axis.line = element_line(lineend = "round"),
    axis.text = element_text(color = "black"),
    axis.ticks = element_line(color = "black")
  )+
  coord_cartesian(xlim = c(0,1),
                  ylim = c(0,1),
                  clip = "off")


ggsave(filename = here::here("output", "figures", "Fig2_DC_DR.png"), 
       width = 8, height = 5)
```